Intercoder ICBEdataset_figure_semantic_embeddings

### 2022-03-23 replication_data/rmd/03_aggregation.Rmd

@@

options(warn = -1) #Groups that have no values generate max errors
#This doesn't have this yet
#event_number_int,  event_number_int_condition
rex_paste <- function(x){ paste0(x  %>% trimws() %>% sort() %>% unique() %>% na.omit(), collapse=";")  }
temp_long <- 
  icb_long_clean %>% 
    #filter(crisno==keep_crisis & sentence_number_int_aligned==keep_sentence )  %>% #for debugging
    dplyr::select(email_id,crisno,sentence_number_int_aligned, section, sentence, event_number_int, event_number_int_condition,variable_normalized, value_normalized) %>% 
    dplyr::filter(value_normalized!='' & !is.na(value_normalized)) %>%
    dplyr::group_by(crisno,sentence_number_int_aligned, event_number_int_condition, variable_normalized, value_normalized) %>%
    dplyr::select(-event_number_int) %>%
    dplyr::distinct() %>%
    dplyr::mutate(
      email_id = email_id %>% as.factor(), 
      #crisno= crisno %>% as.factor(),
      #sentence_number_int_aligned= sentence_number_int_aligned %>% as.factor(),
      crisno_sentence_number_int_aligned= paste0(crisno, "_", sentence_number_int_aligned) %>% as.factor(), #make sure these are the only factors or it'll try to criss cross sentence number and crisno etc.
      section= section %>% as.character(),
      sentence= sentence %>% as.character(),
      variable_normalized= variable_normalized %>% as.factor()
    ) %>%
    dplyr::select(email_id , crisno,sentence_number_int_aligned, event_number_int_condition, variable_normalized, value_normalized) %>%
    distinct() %>%
    dplyr::group_by(email_id , crisno,sentence_number_int_aligned, event_number_int_condition, 
             variable_normalized,.drop=FALSE) %>% 
      #This drop false means every coder gets a '' whether they filled in that variable or not #section, sentence going to have to ignore the sentence text for now
      #You'll get a warning about adding it as a group
      summarise(value_normalized=value_normalized %>% rex_paste() ) %>%
    #that will add in dead ones now we want to split them back out
    dplyr::mutate(value_normalized = value_normalized %>% stringr::str_replace_all(";;",";")) %>% 
    dplyr::mutate(value_normalized = strsplit(as.character(value_normalized), ";", fixed=T)) %>% 
    tidyr::unnest(value_normalized) %>%
    dplyr::distinct() %>%
    #add a confidence score for every token
    dplyr::mutate(confidence=ifelse(variable_normalized=='raterconfidence', value_normalized, NA)) %>%

    dplyr::mutate(confidence=confidence %>% factor(levels=c('none','low','high','complete')) %>% as.numeric() ) %>%
    dplyr::mutate(confidence_expert= email_id %>% stringr::str_detect('expert') %>% ifelse(confidence,NA) %>% max(na.rm=T) ) %>%
    dplyr::mutate(confidence_undergrad= email_id %>% stringr::str_detect('undergrad') %>% ifelse(confidence,NA) %>% max(na.rm=T) ) %>%

@@

>>>>

```{r}
my.max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA)

@@

options(warn = -1) #Groups that have no values generate max errors
#This doesn't have this yet
#event_number_int,  event_number_int_condition
rex_paste <- function(x){ paste0(x  %>% trimws() %>% sort() %>% unique() %>% na.omit(), collapse=";")  }

for(crisis in unique(icb_long_clean$crisno)){
  print(crisis)

  temp_long_partial <- 
    icb_long_clean %>% 
      dplyr::filter(crisno == crisis) %>%
      #filter(crisno==keep_crisis & sentence_number_int_aligned==keep_sentence )  %>% #for debugging
      dplyr::select(email_id,crisno,sentence_number_int_aligned, section, sentence, event_number_int, event_number_int_condition,variable_normalized, value_normalized) %>% 
      dplyr::filter(value_normalized!='' & !is.na(value_normalized)) %>%
      dplyr::group_by(crisno,sentence_number_int_aligned, event_number_int_condition, variable_normalized, value_normalized) %>%
      dplyr::select(-event_number_int) %>%
      dplyr::distinct() %>%
      dplyr::mutate(
        email_id = email_id %>% as.factor(), 
        #crisno= crisno %>% as.factor(),
        #sentence_number_int_aligned= sentence_number_int_aligned %>% as.factor(),
        crisno_sentence_number_int_aligned= paste0(crisno, "_", sentence_number_int_aligned) %>% as.factor(), #make sure these are the only factors or it'll try to criss cross sentence number and crisno etc.
        section= section %>% as.character(),
        sentence= sentence %>% as.character(),
        variable_normalized= variable_normalized %>% as.factor()
      )  %>%
      dplyr::select(email_id , crisno,sentence_number_int_aligned, event_number_int_condition, variable_normalized, value_normalized) %>%
      distinct() %>%
      dplyr::group_by(email_id , crisno, sentence_number_int_aligned, event_number_int_condition, 
               variable_normalized,.drop=FALSE) %>% 
        #This drop false means every coder gets a '' whether they filled in that variable or not #section, sentence going to have to ignore the sentence text for now
        #You'll get a warning about adding it as a group
        summarise(value_normalized=value_normalized %>% rex_paste() ) %>%
      #that will add in dead ones now we want to split them back out
      dplyr::mutate(value_normalized = value_normalized %>% stringr::str_replace_all(";;",";")) %>% 
      dplyr::mutate(value_normalized = strsplit(as.character(value_normalized), ";", fixed=T)) %>% 
      tidyr::unnest(value_normalized) %>%
      dplyr::distinct() %>%
      #add a confidence score for every token
      dplyr::mutate(confidence=ifelse(variable_normalized=='raterconfidence', value_normalized, NA)) %>%
      dplyr::mutate(confidence=confidence %>% factor(levels=c('none','low','high','complete')) %>% as.numeric() ) %>%
      dplyr::mutate(confidence_expert= email_id %>% stringr::str_detect('expert') %>% ifelse(confidence,NA) %>% my.max() ) %>%
      dplyr::mutate(confidence_undergrad= email_id %>% stringr::str_detect('undergrad') %>% ifelse(confidence,NA) %>% my.max() )  %>%

@@

 if(exists("temp_long")){
   temp_long <- dplyr::bind_rows(temp_long, temp_long_partial)
   }else{
     temp_long <- temp_long_partial
   }
}
```

### 2022-03-23 replication_paper/rmd/07_ICBEdataset_figure_semantic_embeddings.Rmd

To set python, better to set it in ~/.Renviron. `use_python()` doesn't seem to work as desired. See [stackoverflow](https://stackoverflow.com/questions/50145643/unable-to-change-python-path-in-reticulate/58743111#58743111).

@@

#devtools::install_github("ropenscilabs/umapr")
library(umap)
library(umapr)
embedding <- umap(sentences_embedding)

@@

library(RColorBrewer)
brewer_pal_info_df <- brewer.pal.info %>% dplyr::add_rownames( var = "rowname") %>% dplyr::arrange()
brewer_list <- list() 
for(i in c(11, 1:nrow(brewer_pal_info_df))){
  brewer_list[[as.character(i)]] <- brewer.pal(n=brewer_pal_info_df$maxcolors[i], brewer_pal_info_df$rowname[i])
}
posible_colors <- unlist(brewer_list) %>% unique(brewer_list) #%>% sample()

value_normalized=codings_long_agreed_embeded_subset$value_normalized %>% unique() %>% as.factor()
value_normalized_colors = posible_colors[1:length(value_normalized)] #ok so pick a bewer pallet with lots of values and then assign it to specific values
names(value_normalized_colors) = levels(value_normalized)

library(ggplot2)
p_semantic_embeddings <-
      codings_long_agreed_embeded_subset %>%
      ggplot(aes(x=UMAP1,y=UMAP2, color=value_normalized )) +
        geom_point(size=0.5, alpha=0.5) + 
        geom_text(data=sampled, aes(x=UMAP1, y=UMAP2, label=crisno_sent), color='black',size=1.5) +

        facet_wrap( ~ variable_normalized) + 
        theme_bw() +
        theme(legend.position = "none") + 
        geom_label(data=value_means,
                   aes(x=UMAP1,y=UMAP2, fill = value_normalized, label=value_normalized, size=50),inherit.aes=F, # #variable_normalized_count
                   colour = "white", fontface = "bold", label.padding = unit(0.1, "lines")) + 
        labs(
            title = "Variation in Tags by Semantic Embeddings of Source Sentences",
            subtitle = "Sentences Embedding Sentence-BERT, UMAP Projection ",
            caption = ""
        ) +
       scale_color_manual(values = value_normalized_colors) +
       scale_fill_manual(values = value_normalized_colors)

>>>>

To set python, set it in ~/.Renviron. `use_python()` doesn't seem to work as desired. See [stackoverflow](https://stackoverflow.com/questions/50145643/unable-to-change-python-path-in-reticulate/58743111#58743111).

@@

embedding <- umapr::umap(sentences_embeddings)

@@

value_means <- value_means %>% 
  #dplyr::mutate(UMAP1 = UMAP1+1, UMAP2 = UMAP2+1) %>% 
  dplyr::mutate(UMAP1=replace(UMAP1, value_normalized=='invasion/occupation', as.numeric(value_means[value_normalized=='invasion/occupation','UMAP1'])+.5)) %>%
  dplyr::mutate(UMAP1=replace(UMAP1, value_normalized=='meeting', as.numeric(value_means[value_normalized=='meeting','UMAP1'])-.3)) %>%
  dplyr::mutate(UMAP2=replace(UMAP2, value_normalized=='invasion/occupation', as.numeric(value_means[value_normalized=='invasion/occupation','UMAP2'])-.7)) %>%
  dplyr::mutate(UMAP2=replace(UMAP2, value_normalized=='meeting', as.numeric(value_means[value_normalized=='meeting','UMAP2'])-.3))

```{r}
do_leaf_all <- codings_long_agreed_embeded_subset %>% dplyr::filter(variable_normalized=="leaf_do_original") %>% dplyr::count(value_normalized) %>% arrange(desc(n)) 
do_leaf_grey <- do_leaf_all %>% dplyr::slice_tail(n=(nrow(do_leaf_all)-10))

value_normalized = codings_long_agreed_embeded_subset %>% count(variable_normalized, value_normalized) %>% arrange(variable_normalized, desc(n)) %>% pull(value_normalized) %>% factor()

# safe_colorblind_palette <- c(#"#88CCEE", drop light blue
#   "#CC6677", 
#   #"#DDCC77", drop yellow
#   "#117733", "#332288", "#AA4499", "#44AA99", "#999933", "#882255", "#661100", "#6699CC"#, "#888888" drop grey
#   )

# https://jacksonlab.agronomy.wisc.edu/2016/05/23/15-level-colorblind-friendly-palette/
  pal <- c(#"#000000",
  "#6db6ff","#004949","#ff6db6", 
  #"#ffb6db", "#24ff24",
 "#db6d00","#b66dff","#009292",
 #"#b6dbff",
 "#920000","#924900","#006ddb","#490092"
 #,"#ffff6d"
 )

value_normalized_colors = rep(pal, ceiling(length(value_normalized)/length(pal)))
names(value_normalized_colors) = as.character(value_normalized)

value_normalized_colors[names(value_normalized_colors) %in% do_leaf_grey$value_normalized] <- "#c8c8c8"

```{r}
value_means <- value_means %>% 
                dplyr::mutate(label=ifelse(
                variable_normalized=="leaf_do_original" & value_normalized %in% do_leaf_grey$value_normalized, 0, 1))

library(ggplot2)
p_semantic_embeddings <-
      codings_long_agreed_embeded_subset %>%
      ggplot(aes(x=UMAP1,y=UMAP2, color=value_normalized )) +
        geom_point(size=0.5, alpha=0.5) + 
        geom_text(data=sampled, aes(x=UMAP1, y=UMAP2, label=crisno_sent), color='black',size=1.5) + 
        facet_wrap( ~ variable_normalized) + 
        theme_bw() +
        theme(legend.position = "none") + 
        geom_label(data=subset(value_means, label==1),
                   aes(x=UMAP1,y=UMAP2, fill = value_normalized, 
                       label=value_normalized, size=50), size = 3, inherit.aes=F, # #variable_normalized_count
                   colour = "white", fontface = "bold", label.padding = unit(0.1, "lines")) + 
        labs(
            title = "Variation in Tags by Semantic Embeddings of Source Sentences",
            subtitle = "Sentences Embedding Sentence-BERT, UMAP Projection ",
            caption = ""
        ) +
       scale_color_manual(values = value_normalized_colors) +
       scale_fill_manual(values = value_normalized_colors)
